---
title: BIOS 731 Homework 3
author: Liam Hanson
format: html
---
### Github Repository:
https://github.com/lhanson01/BIOS_731_HW3

## Problem 1: Newton's Method

The derived gradient and hessian of the log-likelihood are 

$\nabla l(\beta) = [\sum_{i=1}^nX_{ij}(y_i -\pi_i)]_{j \in {1,...,p}}$

and

$\nabla^2l(\beta) = [-\sum_{i=1}^nX_{ij}X_{ik}\pi_{i}]_{j,k \in {1,...,p}}$

where

$$
\pi_i = \frac{e^{X_i^T\beta}}{1+e^{X_i^T\beta}}.
$$ The newton update is $\beta_{t+1} = \beta_t + [\nabla^2l(\beta)]^{-1}\nabla l(\beta)$.

The log-likelihood $l(\beta)$ is convex because the the negative of the hessian $\nabla^2l(\beta)$ is positive semi-definite. See attached derivations for proof.

## Problem 2: MM

See attached derivations. Was not able to implement.

## Problem 3: Simulation Study

```{r}
#| include: false

here::i_am("analysis/HW3_final_report.qmd")
source(here::here("source","01_newton.R"))
source(here::here("source","03_optim.R"))
source(here::here("source","run_simulations.R"))
```

```{r}
plot(newton_ll_history, type = "l", main = "Newton log-likelihood", ylab = "Log LL",
     xlab = "Iterations")

plot(glm_ll_hist, type = "l", main = "GLM log-likelihood", ylab = "Log LL",
     xlab = "Iterations")

plot(obj_history, type = "l", main = "Optim log-likelihood", ylab = "Log LL",
     xlab = "Iterations")

barplot(c(t_glm, t_newton, t_optim), main = "Computation time", xlab = "Method", ylab = "Time Elapsed in seconds", names.arg = c("GLM", "Newton", "Optim"))

print(results_frame)
```
## Problem 4: EM

See derivations attachment. The fitted $\lambda$ value using EM was 130.18. Percentile bootstrap confidence levels were obtained at $\alpha = 0.005$ giving interval [104, 160].

Survival package function survreg also gave an estimated $\lambda$ of 130.18. The estimated 95% confidence interval was [109, 155]. 


